---
title: "Notes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Introductory Notes
====================================

This is not a textbook and it is not intend to be. 
Rather, it is short summary that emphases messages of lectures, 
gives corresponding references and reminds tasks (if any). 
During the course this note will be updated regularly.

### Students assesments

The final score are going to be estimated according a formula given below. 
Give your self a time to understand it: 

```{r, eval=FALSE}
round(sum(min(6, sum(tasks, midExam, extraPoints)), goldenPoints, finalExam))
```

The variables are explained in more details:

* `tasks`: during the course you are going to have quite a number of tasks. 
I believe this is the main part of the course. The maximum sum of points is 4, but
please note, that huge (unlimited) number of extra points can be received, 
if you demonstrate extra efforts in the tasks. In addition, please note that the `tasks` 
variable can be positive only if you supplied all the tasks that was required. 
If at least on of obligatory tasks are missing, the sum of total points are going to be 0.
Thus, do all the tasks, event if the deadline is far over.
* `midExam`: in the middle of semester you are going to have mid exam. The possible outcome is 2 points.
* `extraPoints`: extra points are going to serve as system of encouraging. You can get extra points in many situations: 
taking challenges, answering unexpected question in lecture, giving good question and so on. 
Please note, that `extraPoints` can be negative as well. This could happen in the case of cheating (or something).
* `goldenPoints`: golden points are very similar to extra points, but to get those, 
you will have to put extra efforts in it. How to get them will be steaded in course.
* `finalExam` after the semester you are going to have final exam. The total points of the exam is 4.

**Cultural aspects of the course:**

1. During the course students are encourage to collaborate with each other. 
You can freely share thought, ideas and event source codes. 
In addition students can use all the material that can be found in the Internet or textbooks.
Using the materials made by others are absolutely legal and have no negative side effect on your assesment.
But there are two conditions that are strictly mandatory:
  * You have to give all the credits to original source. 
  Please write an explicit note with links and give some comments. 
  * You have to make sure you understand all the details of the code (or ideas) that you are using. 
  You can be asked to comment it in great details. So, please use the material that you fully understand.

2. I am promise to do my best in answering all your question. You can ask freely during the lectures.
Nevertheless, in practical assignments, your tasks is to try to find the answer by your self. 
So, in practical tasks, before asking any question please be prepare to report that 
efforts you have made to get the answer?
For example: what has the Google query that you have used? That function you have already tried to apply?  


3. Reading and assessing other students rapports is an important part of your learning. 
Please be honest to your colleagues and give the best insights. On the other hand, 
you should appreciate all the criticism that you get and accept it as important part of your learning.

4. Any initiative and communication is highly encouraged. 
Thus, If you have any ideas how to make course more productive or interesting, please let me know. 

5. The tasks are always open for deeper analysis. If you
have idea how to reveal topic in more detail please do it - it probably will get you some extra points.
On the other hand, if you are not satisfied with the assignment that you get, 
you are free to give your reasons and negotiate.



Chapter 1. Technical tools
=====================================

## Technical requirements

In the course you are going to use tree main software:

* R - environment for statistical computing and graphics
* RStudio - a set of integrated tools designed to help you be more productive with R.
* Git - distributed version control system 


I assume you already know R and how to install it. If not check the links below:

* [Installing R on Windows](https://www.youtube.com/watch?v=mfGFv-iB724&feature=youtu.be)
* [Installing R on Mac](https://www.youtube.com/watch?v=Icawuhf0Yqo&feature=youtu.be)
* [Installing R on Ubuntu (linux) ](https://www.youtube.com/watch?v=P8wx4HY9me0)

In addition to R, it’s highly recommended that you install RStudio, 
which will make your experience with R much more enjoyable.
If you need to install RStudio, you can do so [here](https://www.rstudio.com/products/rstudio/download/).
Select the appropriate installer for your operating system.

Moreover, Git is also quite an important software. 
Git is a free and open source distributed version control system.
It helps you not to get lost in versions of your project. 
In addition it is much easier to merge contributions among 
all team members and always have the newest version of the project.
This is a system that professionals uses quite a lot (please admire it).
Please visit [Git home](https://git-scm.com/) and install it in your system.
Originally Git is command line program, on the other hand, Git might be used with
graphical interface (GUI). Visit [Gits guis page](https://git-scm.com/downloads/guis) 
to select the one that looks best for you.

## Task 1. Rmd in Github

The goal of your first task is to get familiar Git. 

1. Firstly, you have to sign up for [Github account](https://github.com/) and then create a repository under the name `Econometrics1`. 

2. Write an email to me (Vygantas.Butkus@mii.vu.lt) with subject "[PE1] Link to Github". And give the link to your Github account in the body.

Hint: A [short learning course on git](https://try.github.io/levels/1/challenges/1 could be useful.
Please try it if you are not familial with Git.
The deadline of this task is 2016-09-09 23:59.



## Task 2. Swirl course

Please visit [swirl](http://swirlstats.com/) website. 
Swirl is a frameworks for interactive R learning.
Your task is to complete R programming course that can be find
in my [GitHub account](https://github.com/1vbutkus/R_Programming).

For a first time, you have to install `swirl` package and the course material. 
Run R commands (be patient, this can take a while):
```{r, eval=FALSE}
install.packages(c('swirl', 'base64enc'))
library("swirl")
install_course_github("1vbutkus", "R_Programming")
```


To run a course you should type:
```{r, eval=FALSE}
library("swirl")
swirl()
```

Follow the instructions and select the "R Programming" course.
Course contains 16 lessons. You have to do all of them. After 
each lesson you have to submit log code that will be generated by swirl.
In addition, save the log code (just in case).

Paskaita 1.
==========================================

R atsitikinių dydžių generavimas
```{r}
x = rnorm(100000)
mean(x<=0)
hist(x)
```


Lošimo kauliukų ekspermentas
```{r}
experment <- function() sum(sample(1:6, 3, replace = TRUE)) > sum(sample(1:6, 2, replace = TRUE))
res = replicate(1000, experment())
mean(res)

```


Paskaita 2.
==========================================

Pavyzdys paskaitoje. Nubrėšime grafiką, kaip įvykio pasirodymo vidurkis artėja į teorinę tikimybę.


```{r}
n = 1000
x = sample(0:1,n, replace = TRUE) 

# skaiciuojame slenkancia suma - tai neefektyvus dudas, bet jis parodo skaiciavimo logika
S_bad = x * 0
S_bad[1] = x[1]
for(i in 2:n){
  S_bad[i] = S_bad[i-1] + x[i] 
}

# efektyvus budas
S = cumsum(x)

# isitikinkime, kad abu budai duoda ta pati rezultata
all(S == S_bad)

# nubrezkime grafika
plot(S / 1:n, type="l", ylab = "S / n", main="Santykinio dažnio arėjimas į teorinę tikimybę")
abline(h=0.5, col=2)
```





## Task 3. Rmd in Github

The goal of your first task is to get familiar with R, Rmd files and Git. 

1. I assume you already have github account with fresh repository `Econometrics1`.
If not, go to task1.

2. Go to the Google and find some awesome R features, packages, programs, graphs or something.
Pick anything that you think is worth noting. 

2. In `Econometrics1` (in your account) create a subdirectory `task3` and in this subdirectory
create Rmd file with title `task3.Rmd`. Then write a very short summary about the topic you have choose.
Please be polite and give the link to original material.
Try to put some R code snippets and visualization that describes the main essence of your note.
Try to customize something (if possible). And finally make sure that the file can be knited.

3. Knit Rmd file. Make sure everything is OK.

4. Apply `add` command to include all related files. That include Rmd file, 
all pictures files and the final output file. Commit it to Git and push to Gith.
Please note, that RStudio has small GUI for Git. 
It is available if you create a project (instead of plain Rmd file).

5. Please check in your github account if all files are uploaded.

Hint: A [short learning course on git](https://try.github.io/levels/1/challenges/1 could be useful.
Please try it if you are not familial with Git.

Please note that those who demonstrate the greatest quality in their task will have actual opportunity 
to get their first extra points. 

The deadline of this task is 2016-09-17 23:59.

## Task 4. Reading

Read [@Kerns2011a, ch. 4] and make sure you understand everything.

References
============================================================
---
references:
- id: Kerns2011a
  title: Introduction to Probability and Statistics Using R
  author:
  - family: Kerns
    given: G. Jay
  issued:
    year: 2011

---
